---
title: "Building the Hawaii Coastal Sites and <br> Characteristics Dataset"
author: "Alemarie Ceria"
output: 
  html_document:
    toc: true
---

<style type="text/css">

h1.title {
  font-size: 38px;
  color: Black;
  font-weight:bold;
  text-align: center;
}
h4.author { /* Header 4 - and the author and data headers use this too  */
    font-size: 18px;
  color: Black;
  text-align: center;
}
h4.date { /* Header 4 - and the author and data headers use this too  */
  font-size: 18px;
  color: Black;
  text-align: center;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

***

## Libraries

```{r Libraries, message=FALSE, warning=FALSE}
# For finding path to directory
library(here)

# For data tidying
library(tidyverse)

# For visualizations
library(rmarkdown)
library(leaflet)
```

## Data Sources

1. Geographic Names Dataset
    - The dataset comes from the Hawaii Statewide GIS Program website and can be downloaded [here](https://geoportal.hawaii.gov/datasets/HiStateGIS::gnis-geographic-names/about) 
    - Metadata can be found [here](https://files.hawaii.gov/dbedt/op/gis/data/geonames.pdf)
2. Webscraped data from [To-Hawaii.com](https://www.to-hawaii.com/maui/beaches/) blog
    - See `code/data_collection_methods.Rmd` file for step-by-step process
3. Google Maps Places API - [Nearby Search](https://developers.google.com/maps/documentation/places/web-service/search-nearby)
    - See `code/data_collection_methods.Rmd` file for step-by-step process

## Data tidying

### Geographic Names Dataset

#### Define a function to perform selecting, renaming and filtering

```{r select_rename_filter() Function, message=FALSE, warning=FALSE}
# 
select_rename_filter <- function(data) {
  data %>%
    select(feature_name, feature_class, county_name, map_name, X, Y) %>%
    rename(lat = Y, lon = X, site_name = feature_name, 
           class = feature_class, county = county_name) %>%
    # Only keep beach parks
    filter(class %in% c("Bay", "Beach", "Harbor") |
           str_detect(site_name, regex("\\bbeach park\\b", ignore_case = TRUE)))
}

# Validation
# Parks total: 513 (With other classes: 1109),
# select_rename_filter <- function(data) {
#   data %>%
#     select(feature_name, feature_class, county_name, map_name, X, Y) %>%
#     rename(lat = Y, lon = X, site_name = feature_name, class = feature_class) %>%
#     filter(class %in% c("Bay", "Beach", "Harbor", "Park"))
# }
# Beach parks total: 105 (With other classes: 701)
# geographic_names_df %>%
#   filter(class == 'Park' & str_detect(site_name, regex("\\bbeach park\\b", ignore_case = TRUE)))
```

#### Read in dataset

```{r Geographic Names Raw Dataset, message=FALSE, warning=FALSE}
geographic_names <- read.csv(here('data/raw/geographic_names_hi_gis.csv'))
```

#### Create the following source objects

```{r Source Objects, message=FALSE, warning=FALSE}
# Maui island bounding box coordinates
min_lon <- -156.712611
max_lon <- -155.957301
min_lat <- 20.568013
max_lat <- 21.042995

# Kahoolawe site that will get removed by bounding box -- Add back in later
kaulana_beach <- geographic_names %>% 
  filter(feature_name == 'Kaulana Beach') %>% 
  select_rename_filter(.)
```

#### Tidy dataset

```{r Tidy Geographic Names Dataset, message=FALSE, warning=FALSE}
geographic_names_df <- geographic_names %>% 
  select_rename_filter(.) %>%
  # Remove Maui Island sites -- Derek already did it
  filter(lon < min_lon | lon > max_lon | lat < min_lat | lat > max_lat) %>% 
  # Add Kaulana Beach back in
  bind_rows(., kaulana_beach) %>% 
  # Remove sites not a part of the Main Hawaiian Islands (Adams Bay and West Cove)
  filter(site_name != 'Adams Bay' & site_name != 'West Cove' & site_name != 'Shark Bay') %>%
  # Standardize site_name
  mutate(site_name = stringi::stri_trans_general(site_name, "Latin-ASCII")) %>%
  mutate(site_name = stringr::str_trim(site_name)) %>% 
  # Remove "'" and "(historical)"
  mutate(site_name = stringr::str_remove(site_name, "'")) %>%
  mutate(site_name = stringr::str_remove_all(site_name, "\\(.*\\)")) %>% 
  mutate_if(is.character, as.factor)
  # mutate(site_name = stringi::stri_trans_tolower(site_name))
```

#### Table

```{r Tidied Geographic Names Dataset, message=FALSE, warning=FALSE}
paged_table(geographic_names_df)
```

#### Interactive map

Click site to see site name 

```{r Tidied Geographic Names Map, fig.height=5, fig.width=9.5, message=FALSE, warning=FALSE}
# Creating a qualitative color scale:
class_colors <- colorFactor(c('#d7191c', '#fdae61', '#abd9e9', '#2c7bb6'),
                             domain = c('Bay', 'Beach', 'Harbor', 'Park'))

leaflet(geographic_names_df) %>% 
  addTiles() %>%
  setView(-157.56868557054815, 20.455, zoom = 7) %>%
  addCircleMarkers(~lon, ~lat, popup = geographic_names_df$site_name, 
                   weight = 3, color = ~class_colors(class), 
                   radius = 4, stroke = F, fillOpacity = 0.75) %>% 
addLegend(colors = c('#d7191c', '#fdae61', '#abd9e9', '#2c7bb6'), labels=c('Bay', 'Beach', 'Harbor', 'Park'), "bottomleft", title="Class", opacity = 0.9)
```

***

## Data Dictionary

```{r Data Dictionary, echo=FALSE, message=FALSE, warning=FALSE}
data_dictionary <- read.csv(here('documentation/coastal_sites_characteristics_data_dictionary.csv'))

paged_table(data_dictionary)
```

```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
# For future reference: You can also use the package fuzzyjoin to join the datasets based on approximate matching of site name

library(fuzzyjoin)
fuzzy_left_join(geographic_names_df, other_data_source, by = c("site_name" = "site_name"), match_fun = list(`==`, stringdist_inner_join))
```

