---
title: "Building the Hawaii Coastal Sites and <br> Characteristics Dataset"
author: "Alemarie Ceria"
output: 
  html_document:
    toc: true
---

<style type="text/css">

h1.title {
  font-size: 38px;
  color: Black;
  font-weight:bold;
  text-align: center;
}
h4.author { /* Header 4 - and the author and data headers use this too  */
    font-size: 18px;
  color: Black;
  text-align: center;
}
h4.date { /* Header 4 - and the author and data headers use this too  */
  font-size: 18px;
  color: Black;
  text-align: center;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

***

## Libraries

```{r Libraries, message=FALSE, warning=FALSE}
# For finding path to directory
library(here)

# For data tidying
library(tidyverse)

# For visualizations
library(rmarkdown)
library(leaflet)
library(sp)
```

## Data Sources

1. [Geographic Names Dataset](https://geoportal.hawaii.gov/datasets/HiStateGIS::gnis-geographic-names/about)
    - This dataset will be used as an extensive list of coastal sites in Main Hawaiian Islands (MHI)
      and will be filtered according to the following datasets which will tell us where people actually
      recreate
2. Webscraped data from [To-Hawaii.com](https://www.to-hawaii.com/maui/beaches/) blog
    - [About](https://www.to-hawaii.com/about-us.php) To-Hawaii.com
      - TLDR:
        - Created in 2006 by two residents to serve as an information portal for visitors and those interested in Hawaii
        - Provides in-depth and unbiased information about the Hawaiian Islands and its culture, history, language, art, and challenges
        - Covers all six MHI and offers detailed information on each island's unique experiences and attractions
        - Mission is to provide the most comprehensive information about Hawaii available online
    - This dataset will be used to gather information on: name of coastal sites, classification, sand composition and type, facilities, activities, marine mammal observation, and access/parking fees if available
    - See `code/data_collection/webscrape_to_hawaii_sites.py` file for python script
3. Google Maps Places API - [Nearby Search](https://developers.google.com/maps/documentation/places/web-service/search-nearby)
    - This dataset will be used to gather names of coastal sites based on total number of reviews (i.e., The more reviews, the more popular the site), classification, sand composition and type, facilities, activities, marine mammal observation, and access/parking fees if available
    - See `code/data_collection/request_places_api_sites.py` file for python script

## Data tidying

### Geographic Names Dataset

#### Define a function to perform selecting, renaming and filtering

```{r select_rename_filter() Function, message=FALSE, warning=FALSE}
select_rename_filter <- function(data) {
  data %>%
    select(feature_name, feature_class, county_name, map_name, Y, X) %>%
    rename(name = feature_name, 
           classification = feature_class, 
           county = county_name,
           lat = Y, lon = X) %>%
    # Only keep beach parks
    filter(classification %in% c("Bay", "Beach", "Harbor") |
           str_detect(name, regex("\\bbeach park\\b", ignore_case = TRUE)))
}

# Validation
# Parks total: 513 (With other classes: 1109),
# select_rename_filter <- function(data) {
#   data %>%
#     select(feature_name, feature_class, county_name, map_name, X, Y) %>%
#     rename(lat = Y, lon = X, site_name = feature_name, class = feature_class) %>%
#     filter(class %in% c("Bay", "Beach", "Harbor", "Park"))
# }
# Beach parks total: 105 (With other classes: 701)
# geographic_names_df %>%
#   filter(class == 'Park' & str_detect(site_name, regex("\\bbeach park\\b", ignore_case = TRUE)))
```

#### Read in dataset

```{r Geographic Names Raw Dataset, message=FALSE, warning=FALSE}
geographic_names <- read.csv(here('data/raw/usgs_geographic_names_sites.csv'))
```

#### Create the following source objects

```{r Source Objects, message=FALSE, warning=FALSE}
# Maui island bounding box coordinates
min_lon <- -156.712611
max_lon <- -155.957301
min_lat <- 20.568013
max_lat <- 21.042995

# Kahoolawe site that will get removed by bounding box -- Add back in later
kaulana_beach <- geographic_names %>% 
  filter(feature_name == 'Kaulana Beach') %>% 
  select_rename_filter(.)
```

#### Tidy dataset

```{r Tidy Geographic Names Dataset, message=FALSE, warning=FALSE}
geographic_names_df <- geographic_names %>% 
  select_rename_filter(.) %>%
  # Remove Maui Island sites -- Derek already did it
  filter(lon < min_lon | lon > max_lon | lat < min_lat | lat > max_lat) %>% 
  # Add Kaulana Beach back in
  bind_rows(., kaulana_beach) %>% 
  # Remove sites not a part of the Main Hawaiian Islands (Adams Bay and West Cove)
  filter(name != 'Adams Bay' & name != 'West Cove' & name != 'Shark Bay') %>%
  # Standardize site_name
  mutate(name = stringi::stri_trans_general(name, "Latin-ASCII")) %>%
  mutate(name = stringr::str_trim(name)) %>% 
  # Remove "'" and "(historical)"
  mutate(name = stringr::str_remove(name, "'")) %>%
  mutate(name = stringr::str_remove_all(name, "\\(.*\\)")) %>% 
  mutate_if(is.character, as.factor)
  # mutate(site_name = stringi::stri_trans_tolower(name))
```

#### Table

```{r Tidied Geographic Names Dataset, message=FALSE, warning=FALSE}
paged_table(geographic_names_df)
```

#### Interactive map

Click site to see site name 

```{r Tidied Geographic Names Map, fig.height=5, fig.width=9.5, message=FALSE, warning=FALSE}
# Creating a qualitative color scale:
class_colors <- colorFactor(c('#d7191c', '#fdae61', '#abd9e9', '#2c7bb6'),
                             domain = c('Bay', 'Beach', 'Harbor', 'Park'))

leaflet(geographic_names_df) %>% 
  addTiles() %>%
  setView(-157.56868557054815, 20.455, zoom = 7) %>%
  addCircleMarkers(~lon, ~lat, popup = geographic_names_df$site_name, 
                   weight = 3, color = ~class_colors(class), 
                   radius = 4, stroke = F, fillOpacity = 0.75) %>% 
addLegend(colors = c('#d7191c', '#fdae61', '#abd9e9', '#2c7bb6'), labels=c('Bay', 'Beach', 'Harbor', 'Park'), "bottomleft", title="Class", opacity = 0.9)
```

***

## Data Dictionary

```{r Data Dictionary, echo=FALSE, message=FALSE, warning=FALSE}
data_dictionary <- read.csv(here('documentation/coastal_sites_characteristics_data_dictionary.csv'))

paged_table(data_dictionary)
```

```{r eval=FALSE, message=FALSE, warning=FALSE, include=FALSE}
# For future reference: You can also use the package fuzzyjoin to join the datasets based on approximate matching of site name

library(fuzzyjoin)
fuzzy_left_join(geographic_names_df, other_data_source, by = c("site_name" = "site_name"), match_fun = list(`==`, stringdist_inner_join))
```

